maze-dataset: Maze Generation with Algorithmic Variety and Representational Flexibility

### 2025-04-26 docs/paper/paper.tex
move figure code to external files

\begin{figure} 
  \begin{minipage}{5in}
    \input{diagram/diagram.tikz} 
  \end{minipage}
  \caption{
    Usage of maze-dataset. We create a \texttt{MazeDataset} from a \texttt{MazeDatasetConfig}. This contains \texttt{SolvedMaze} objects which can be converted to and from a variety of formats. Code in the image contains clickable links to \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html}{documentation}. A variety of generated examples can be viewed \href{https://understanding-search.github.io/maze-dataset/examples/maze_examples.html}{here}.
  }
  \label{fig:diagram}
\end{figure}

@@

\begin{figure}[H]
  \centering
  \begin{tabular}{p{1.5in} p{1.5in} p{1.5in}} 
    \hline \\[.5em]
    % algorithms
    \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze.as_ascii}{\texttt{as\_ascii()}}
    & \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze.as_pixels}{\texttt{as\_pixels()}}
    & \href{https://understanding-search.github.io/maze-dataset/maze_dataset/plotting.html\#MazePlot}{\texttt{MazePlot()}} \\[.5em]
    % descriptions
      Simple text format for displaying mazes, useful for debugging in a terminal environment.
      & \texttt{numpy} array of \texttt{dtype=uint8} and shape \texttt{(height, width, 3)}. The last dimension is RGB color.
      & feature-rich plotting utility with support for multiple paths, heatmaps over positions, and more. \\[1em]
    \hline \\
    % examples
      \multicolumn{1}{c}{\begin{minipage}[b]{1.6in}
        \setlength{\baselineskip}{0.9em}
        \input{figures/outputs-ascii-colored.tex} 
      \end{minipage}}
      & \multicolumn{1}{c}{
        \includegraphics[width=0.25\textwidth]{figures/outputs-pixels.pdf}
      }
      & \multicolumn{1}{c}{
        \includegraphics[width=0.27\textwidth, trim={0 0.8cm -.3cm, -.5cm}, clip]{figures/outputs-mazeplot.pdf}
      } \\[1em]

    \hline \\
  \end{tabular}
  \caption{Various output formats. Top row (left to right): ASCII diagram, rasterized pixel grid, and advanced display tool.}
  \label{fig:output-fmts}
\end{figure}

@@

\begin{figure} 
  \centering
  \begin{minipage}{5in}
    \footnotesize
    \input{figures/outputs-tokens-colored.tex}
  \end{minipage}
  \caption{
    Example text output format with token regions highlighted.
    \colorbox[RGB]{ 217,210,233 }{Adjacency list}: text representation of the graph,
    \colorbox[RGB]{ 217,234,211 }{Origin}: starting coordinate,
    \colorbox[RGB]{ 234,209,220 }{Target}: ending coordinate,
    \colorbox[RGB]{ 207,226,243 }{Path}: maze solution sequence
  }
  \label{fig:token-regions}
\end{figure}

@@

\begin{figure}
    \centering
    \input{figures/TokenizerElement_structure.tikz}
    \caption{Nested internal structure of \texttt{\_TokenizerElement} objects inside a typical \texttt{MazeTokenizerModular}.}
\end{figure}

@@

\begin{table}[H]
\centering
\begin{tabular}{|ll|r|rrr|}
  \hline
  maze\_ctor
          & keyword args           & all sizes 
                                              & \shortstack{small \\ $g \leq 10$} 
                                                         & \shortstack{medium \\ $g \in (10, 32]$} 
                                                                    & \shortstack{large \\ $g > 32$} \\
  \hline\hline
  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_dfs}{dfs}
          &                        &   28.0   &    2.8   &   20.3   &  131.8   \\
  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_dfs}{dfs}
          & accessible\_cells=20   &    2.3   &    2.2   &    2.4   &    2.2   \\
  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_dfs}{dfs}
          & do\_forks=False        &    2.7   &    2.2   &    3.1   &    3.5   \\
  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_dfs}{dfs}
          & max\_tree\_depth=0.5   &    2.5   &    2.0   &    2.7   &    4.0   \\
  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_dfs_percolation}{dfs\_percolation}
          & p=0.1                  &   43.9   &    2.8   &   33.9   &  208.0   \\
  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_dfs_percolation}{dfs\_percolation}
          & p=0.4                  &   48.7   &    3.0   &   36.5   &  233.5   \\
  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_kruskal}{kruskal}
          &                        &   12.8   &    1.9   &   10.3   &   55.8   \\
  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_percolation}{percolation}
          & p=1.0                  &   50.2   &    2.6   &   37.2   &  242.5   \\
  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_recursive_division}{recursive\_div}
          &                        &   10.2   &    1.7   &    8.9   &   42.1   \\
  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_wilson}{wilson}
          &                        &  676.5   &    7.8   &  188.6   & 3992.6   \\
  \hline\hline
  mean
          &                        &  559.9   &   13.0   &  223.5   & 3146.9   \\
  median
          &                        &   11.1   &    6.5   &   32.9   &  302.7   \\
  \hline
\end{tabular}
\caption{Generation times for various algorithms and maze sizes. More information can be found on the \docsref{benchmarks}{benchmarks page}.}
\label{tab:benchmarks}
\end{table}

>>>>

\input{figures/tex/fig1_diagram.tex}

@@

\input{figures/tex/fig2_formats.tex}

@@

\input{figures/tex/fig4_tokenfmt.tex}

@@

\input{figures/tex/fig5_mmt.tex}

@@

\input{figures/tex/tab1_benchmarks.tex}
